Comparing clustering parameters with rOpenScPCA

### 2025-12-19 analyses/cell-type-consensus/exploratory-notebooks/07-compare-consensus.Rmd

# combine into a single df and add a column to indiciate old/new 
df_list <- list(old_consensus_df, new_consensus_df) |> 
  purrr::set_names(c("no_scimilarity", "with_scimilarity"))

>>>>

# combine into a single df and add a column to indicate old/new 
df_list <- list(old_consensus_df, new_consensus_df) |> 
  purrr::set_names(c("no_scimilarity", "with_scimilarity"))

### 2025-12-19 analyses/cell-type-ewings/exploratory_analysis/01-marker-gene-tumor-classification.Rmd

```{r, fig.height=10, fig.width=10}
marker_plot_df <- celltype_df |>
  dplyr::mutate(gene_expresed = dplyr::if_else(transformed_gene_expression > 0, TRUE, FALSE)) |>
  dplyr::select(gene_symbol, gene_expresed, sum_classification)

ggplot(marker_plot_df, aes(x = sum_classification, fill = gene_expresed)) +
  geom_bar(stat = "count") +
  facet_grid(rows = vars(gene_symbol)) +
  scale_fill_manual(values = c("grey", "blue")) +
  theme_classic()
```

>>>>

```{r, fig.height=10, fig.width=10}
marker_plot_df <- celltype_df |>
  dplyr::mutate(gene_expressed = dplyr::if_else(transformed_gene_expression > 0, TRUE, FALSE)) |>
  dplyr::select(gene_symbol, gene_expressed, sum_classification)

ggplot(marker_plot_df, aes(x = sum_classification, fill = gene_expressed)) +
  geom_bar(stat = "count") +
  facet_grid(rows = vars(gene_symbol)) +
  scale_fill_manual(values = c("grey", "blue")) +
  theme_classic()
```

### 2025-12-19 analyses/cell-type-ewings/exploratory_analysis/04-infercnv.Rmd

- If we have a known group of normal cells, `InferCNV` can be helpful in identifying cells that could be tumor cells, e.g., cells that have a high number of CNVs. 
- Using `InferCNV` without a group of normal cells does not seem feasible for this particular sample since all tumor cells appear to have the same set of CNVs with little variation. 
- We can define tumor cells by setting a cutoff for the number of CNVs that must be present in a cell, but I'm not sure how transferrable this is to other samples. 
- `InferCNV` can be helpful in looking at specific CNVs that are expected in a tumor type, something that is harder to look at in a tool like `CopyKAT`.

>>>>

- If we have a known group of normal cells, `InferCNV` can be helpful in identifying cells that could be tumor cells, e.g., cells that have a high number of CNVs. 
- Using `InferCNV` without a group of normal cells does not seem feasible for this particular sample since all tumor cells appear to have the same set of CNVs with little variation. 
- We can define tumor cells by setting a cutoff for the number of CNVs that must be present in a cell, but I'm not sure how transferable this is to other samples. 
- `InferCNV` can be helpful in looking at specific CNVs that are expected in a tumor type, something that is harder to look at in a tool like `CopyKAT`.

### 2025-12-19 analyses/cell-type-wilms-tumor-06/notebook/07_combined_annotation_across_samples_exploration.Rmd
caught one more ospl

| first level annotation | second level annotation | selection of the cells | marker genes for validation | CNV validation |
|----|----|----|----|----|
| normal | endothelial | `compartment == "endothelium" & predicted.score > pred.thr` | *VWF* | no CNV |
| normal | immune | `compartment == "immune" & predicted.score > pred.thr` | *PTPRC*, *CD163*, *CD68* | no CNV |
| normal | kidney | `compartment == "fetal_nephron" & predicted.score > pred.thr & cnv_score < cnv.thr.lower` | *CDH1*, *PODXL*, *LTL* | no CNV |
| normal | stroma | `compartment == "stroma" & predicted.score > pred.thr & cnv_score < cnv.thr.upper` | *VIM* | no CNV |
| cancer | stroma | `compartment == "stroma" & cnv_score > cnv.thr.upper` | *VIM* | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| cancer | blastema | `compartment == "fetal_nephron" & cell_type == "mesenchymal cell" & cnv_score > cnv.thr.upper` | *CITED1* | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| cancer | epithelial | `compartment == "fetal_nephron" & cell_type != "mesenchymal cell" & cnv_score > cnv.thr.upper` | *CDH1* | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| unknown | \- | the rest of the cells | \- | \- |

@@

- `pred.thr` = `r params$predicted.celltype.threshold`
- `cnv.thr.lower` = `r params$cnv_threshold_low`
- `cnv.thr.upper` = `r params$cnv_threshold_high`

>>>>

| first level annotation | second level annotation | selection of the cells | marker genes for validation | CNV validation |
|----|----|----|----|----|
| normal | endothelial | `compartment == "endothelium" & predicted.score > pred_thr` | *VWF* | no CNV |
| normal | immune | `compartment == "immune" & predicted.score > pred_thr` | *PTPRC*, *CD163*, *CD68* | no CNV |
| normal | kidney | `compartment == "fetal_nephron" & predicted.score > pred_thr & cnv_score < cnv_threshold_low` | *CDH1*, *PODXL*, *LTL* | no CNV |
| normal | stroma | `compartment == "stroma" & predicted.score > pred_thr & cnv_score < cnv_threshold_upper` | *VIM* | no CNV |
| cancer | stroma | `compartment == "stroma" & cnv_score > cnv_threshold_upper` | *VIM* | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| cancer | blastema | `compartment == "fetal_nephron" & cell_type == "mesenchymal cell" & cnv_score > cnv_threshold_upper` | *CITED1* | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| cancer | epithelial | `compartment == "fetal_nephron" & cell_type != "mesenchymal cell" & cnv_score > cnv_threshold_upper` | *CDH1* | `proportion_cnv_chr: 1, 4, 11, 16, 17, 18` |
| unknown | \- | the rest of the cells | \- | \- |

@@

- `pred_thr` = `r params$predicted.celltype.threshold`
- `cnv_threshold_low` = `r params$cnv_threshold_low`
- `cnv_threshold_upper` = `r params$cnv_threshold_high`

### 2025-12-19 analyses/hello-clusters/02_compare-clustering-parameters.Rmd
fix typo and re-render html

      # calculate neighbhorhood purity, starting from silhouette_df
      rOpenScPCA::calculate_purity(pca_matrix, silhouette_df)
    }
  )

@@

This section presents one coding strategy to calculate and visualize results when varying two clustering parameters.
In particular, we use faceting to help display all information in one plot, by placing nearest neighbor values on the X-axis and faceting by resolution values.
Since silhouette width and neighhorbood purity calculations using generally similar code, we'll just show neighborhood purity here.

>>>>

      # calculate neighborhood purity, starting from silhouette_df
      rOpenScPCA::calculate_purity(pca_matrix, silhouette_df)
    }
  )

@@

This section presents one coding strategy to calculate and visualize results when varying two clustering parameters.
In particular, we use faceting to help display all information in one plot, by placing nearest neighbor values on the X-axis and faceting by resolution values.
Since silhouette width and neighborhood purity calculations using generally similar code, we'll just show neighborhood purity here.
